### 2021-03-05 paper3/paper3.tex
Use include instead of import
This has the advantage that I can compile one chapter at a time.

%%%%%%%%% ICML 2021 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

@@

Algorithms such as classifiers, however, are trained from large amount of labelled data, and are typically trained to optimise \emph{global} metrics such as average classification accuracy.
%
% \begin{figure*}[t]
% \centering
% \includegraphics[width=0.6\textwidth,page=2]{figures/ideal.pdf}
% \caption{Overview of the zero-shot stratification problem. Training dataset (pairs of input data $x$ and class label $y$) will only contain data points that are labelled ``yes'' by the decision policy. This systematic bias might result in a subgroup(s) to have zero labelled data (in the example above, ``married'' subgroup has zero labelled data). The subgrouping information $s$ is unavailable/unlabelled at deployment time, and is only partially labelled at train time.}
% \vspace{-0.5cm}
% \label{fig:censoring}
% \end{figure*}
In many real-world classification tasks, each labelled class consists of multiple semantically distinct subclasses, or subgroups.
For example, the ``dog'' class label can have finer-grained intra-class variations, such as ``dog indoor'' and ``dog outdoor''.
%
This finer-grained subgrouping information is typically unavailable/unlabelled (e.g. \citep{SohDunAngGuetal20,nam2020learning}).
%
The standard training process brings about two inter-connected challenges: a) classifiers often under-perform on important hidden subgroups (\emph{hidden stratification}) \citep{RayDunCarRe20,SohDunAngGuetal20};
and b) \emph{systematic bias}~\citep{kallus2018residual} affects whether or not entire collections of data points appear in the training dataset,
and can make the classifier unprepared for treating those subgroups in the eventual deployment setting
% and can hamper attempts to correct for bias in the eventual deployment data
(\emph{residual bias}). %\cite{kallus2018residual}.

@@

%\begin{wrapfigure}{!t}{0.6\textwidth}
\begin{figure}[t]
\centering
    %\includegraphics[width=0.9\textwidth]{figures/SSL-framework}
    % \includegraphics[width=\textwidth]{figures/SSL-framework-withPred.pdf}
    \includegraphics[width=\textwidth]{figures/zsf_diagram.pdf}
    \caption{%
    The main components involved in our proposed disentangling method, $f_y$ (debiaser) and $h$ (discriminator). The debiaser is trained to produce encodings, $z_y$ of the data that are invariant to the source dataset and thereby the subgroups identifying it. In order to determine whether a bag of encodings originates from the training set or the deployment set, the discriminator performs an attention-weighted aggregation over the bag dimension to model interdependencies between the samples. In the case of Coloured MNIST where {\color{purple}purple} fours constitute the missing subgroup, the discriminator can identify an encoding of a bag from the training set by the absence of such samples so long as colour information is detectable in $z_y$, serving as an error signal for the debiaser.
    }%
    \label{fig:architecture}
\end{figure}%
% \end{figure}
%\end{wrapfigure}
% \hfill
% \begin{figure}
\begin{figure}[t]
    \centering
    % if you want to make changes to the diagram, it is located here:
    % https://drive.google.com/file/d/1M37m4rVj1EGrFJb9QzK32bs8swepiaZ1/view?usp=sharing
    \includegraphics[width=0.5\textwidth]{figures/fdm.pdf}
    \caption{%
    Illustration of distribution matching.
    In this example, the training set is lacking {\color{purple}purple} 4's.
    By enforcing the subspace $z_y$ to have the same distribution for both the training and deployment set, the model is encouraged to learn a representation that is invariant to colour (or $s$ in general.)
    For this to work, it is crucial that the bags be approximately balanced (perfect bags).
    }%
    \label{fig:matching-diagram}
% \end{subfigure}
% \caption{%
% Overview of the disentangling framework with a perfect (balanced) bag as an inductive bias.
% }%
% \label{fig:framework}
\end{figure}

@@

%Semi-supervised learning \cite{ChaSchZie06} can alleviate the issue of unfairness to the \emph{invisibles} by mixing labeled with unlabeled data, which is usually much cheaper to obtain. 
%
We propose to alleviate the issue of bias against missing subgroups by mixing labelled data with unlabelled data that is usually much cheaper to obtain \citep{ChaSchZie06}. 
%
In this paper, we refer to this set of \emph{unlabelled} data as the
deployment set\footnote{In our experiments, we report accuracy and bias metrics on another independent test set instead of on the unlabelled data that is available at training time.} $\gD_{dep} =\{(x_i)\}$.
%_{i=1}^{N^c}
%
This 
%context 
deployment set has overlap with all sources:
\begin{align}
  \gD_{dep} \cap \Omega_{y=y',s=s'} \neq \varnothing\quad\quad\forall y'\in\gY,\forall s'\in\gS~.
\end{align}
%The deployment set is much like the deployment set: 
Importantly, the deployment set has no information about class labels $y$ or the subgroup labels $s$.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=\columnwidth]{figures/celeba_dep_recon.png}
    \caption{%
    Visualization of our method's solutions for the CelebA dataset, with``smiling females'' as the missing subgroup.
    Column 1 shows the original images from $x$ from the deployment set of CelebA.
    Column 2 shows plain reconstructions generated from $x_\textit{recon}=g(f_y(x), f_s(x))$.
    Column 3 shows reconstruction with zeroed-out $z_s$: $g(f_y(x), 0)$, which effectively visualises $z_y$.
    Column 4 shows the result of an analogous process where $z_y$ was zeroed out instead.
    }%
    \label{fig:celeba-recons}
\end{figure}

@@

  \includegraphics[width=\columnwidth]{figures/cmnist_2v4_partial_acc.pdf}
  \includegraphics[width=\columnwidth]{figures/cmnist_2v4_partial_prr.pdf}
%   \includegraphics[width=\columnwidth]{figures/cmnist_2v4_partial_tpr.pdf}
%   \includegraphics[width=\columnwidth]{figures/cmnist_2v4_partial_tnr.pdf}

@@

  \includegraphics[width=\columnwidth]{figures/cmnist_2v4_miss_s_acc.pdf}
  \includegraphics[width=\columnwidth]{figures/cmnist_2v4_miss_s_prr.pdf}
%   \includegraphics[width=\columnwidth]{figures/cmnist_2v4_miss_s_alt_tpr.pdf}
%   \includegraphics[width=\columnwidth]{figures/cmnist_2v4_miss_s_alt_tnr.pdf}

@@

  \includegraphics[width=\columnwidth]{figures/cmnist_3dig_4miss_acc.pdf}
  \includegraphics[width=\columnwidth]{figures/cmnist_3dig_4miss_hgr.pdf}
%   \includegraphics[width=\columnwidth]{figures/cmnist_3dig_4miss_pr.pdf}
%   \includegraphics[width=\columnwidth]{figures/cmnist_3dig_4miss_tpr.pdf}
%   \includegraphics[width=\columnwidth]{figures/cmnist_3dig_4miss_tnr.pdf}

@@

  \includegraphics[width=\columnwidth]{figures/celeba_gender_smiling_acc.pdf}
  \includegraphics[width=\columnwidth]{figures/celeba_gender_smiling_prr.pdf}
%   \includegraphics[width=\columnwidth]{figures/celeba_gender_smiling_tpr.pdf}
%   \includegraphics[width=\columnwidth]{figures/celeba_gender_smiling_tnr.pdf}

@@

\paragraph{Constructing approximately perfect bags via clustering.}
We cluster the data points from the deployment set into $K=\textrm{dim}(\mathcal{Y})\cdot \textrm{dim}(\mathcal{S})$ number of clusters, i.e. the number of data sources $\Omega_{y,s}$. 
%, using a learned embedding $z$.
%
We use the k-means clustering algorithm, and a recently proposed method based on rank statistics \citep{HanRebEhrVedetal20}.
%
The cluster assignments can then be used to evenly stratify the deployment set into perfect batches, to be used by the subsequent disentangling phase.
%
% The objective for clustering can be written as a binary cross-entropy loss, induced by the following binary classification problem -- for any two data points in the deployment set to decide whether or not they belong to the same cluster:
% %
% \begin{align}
%  \mathcal{L}_{\textit{unsup}} &= -\frac{1}{N^2}\sum_{i,j} [s_{ij}\: \text{log} (C(z_i)^\top C(z_j)) \nonumber\\
%  &+ (1-s_{ij}) \: \text{log}(1-C(z_i)^\top C(z_j))].
%  \label{eq:clustering}
%  \end{align}
% %
% Here, $N^2$ is the amount of sample pairs,  $s_{ij}$ are binary labels capturing that similar data points belong to the same group, $s_{ij}=1$, while dissimilar ones do not, $s_{ij}=0$. 
% %
% Since the deployment set is unlabelled, $s_{ij}$ are unknown. 
% %
% We estimate $s_{ij}$ with the robust rank statistics approach (described below) and use the estimates as pseudo-labels in \eqref{eq:clustering}. 
% %
% The dot-product $C(z_i)^\top C(z_j)$ measures the similarity of the two data points $z_i$ and $z_j$ to score whether they share a cluster.
% %
% This is computed on the embedding of the softmax classifier $C: \mathcal{Z} \rightarrow \{1,\ldots, K\}$. 
% %
% In addition to this unsupervised training, the classifier $C$ is also trained on the labeled data using the standard cross-entropy loss.  
% %We train the classifier $C$ on the training data using a standard cross-entropy loss and the labeled sources $\Omega_{y,s}$. 
% %
% This way, the data points in the deployment set that are similar to the known demographics are assigned to the associated clusters, while the dissimilar data points form the new clusters corresponding to the invisible groups.

% The method of rank statistics \citep{YagStrRosLin11,DeaRuzSegShletal13,HanRebEhrVedetal20} avoids comparing the vectors $z_i$ and $z_j$ directly, and instead ranks the values in $z_i$ and in $z_j$ by their magnitudes, and compares the top-$U$ rankings (we found $U = 5$ to perform well in our experiments).
% %
% If the rankings obtained for two unlabeled data points $z_i$ and $z_j$ are the same, they are assigned to the same class, $s_{ij}=1$. 
% %
As a result of clustering, the data points in the deployment set $\gD_\mathit{dep}$ are labelled with cluster assignments 
$\gD_\mathit{dep}=\{(x_i, c_i)\}$, $c_i = C(z_i)$.
%
We balance $\gD_\mathit{dep}$ so that all clusters have equal size to form a perfect bag
% (see fig. \ref{fig:teaser})
, and use it as a supervision signal for the disentangling step.
%
%Note that our method of constructing the perfect dataset based on clustering does not require solving the linear assignment problem of cluster-source association. This is particularly suitable for exploring the setting where invisible demographics are formed of multiple sources.

@@

\input{appendix.tex}

>>>>

\chapter{Paper 3: Learning with Perfect Bags: Addressing Hidden Stratification with Zero Labelled Data}\label{ch:paper3}
%%%%%%%%% ICML 2021 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

@@

Algorithms such as classifiers, however, are trained from large amount of labelled data, and are typically trained to optimise \emph{global} metrics such as average classification accuracy.
%
% \begin{figure*}[t]
% \centering
% \includegraphics[width=0.6\textwidth,page=2]{paper3/figures/ideal.pdf}
% \caption{Overview of the zero-shot stratification problem. Training dataset (pairs of input data $x$ and class label $y$) will only contain data points that are labelled ``yes'' by the decision policy. This systematic bias might result in a subgroup(s) to have zero labelled data (in the example above, ``married'' subgroup has zero labelled data). The subgrouping information $s$ is unavailable/unlabelled at deployment time, and is only partially labelled at train time.}
% \vspace{-0.5cm}
% \label{fig:censoring}
% \end{figure*}
In many real-world classification tasks, each labelled class consists of multiple semantically distinct subclasses, or subgroups.
For example, the ``dog'' class label can have finer-grained intra-class variations, such as ``dog indoor'' and ``dog outdoor''.
%
This finer-grained subgrouping information is typically unavailable/unlabelled (e.g. \citep{SohDunAngGuetal20,nam2020learning}).
%
The standard training process brings about two inter-connected challenges: a) classifiers often under-perform on important hidden subgroups (\emph{hidden stratification}) \citep{RayDunCarRe20,SohDunAngGuetal20};
and b) \emph{systematic bias}~\citep{kallus2018residual} affects whether or not entire collections of data points appear in the training dataset,
and can make the classifier unprepared for treating those subgroups in the eventual deployment setting
% and can hamper attempts to correct for bias in the eventual deployment data
(\emph{residual bias}). %\cite{kallus2018residual}.

@@

%\begin{wrapfigure}{!t}{0.6\textwidth}
\begin{figure}[t]
\centering
    %\includegraphics[width=0.9\textwidth]{paper3/figures/SSL-framework}
    % \includegraphics[width=\textwidth]{paper3/figures/SSL-framework-withPred.pdf}
    \includegraphics[width=\textwidth]{paper3/figures/zsf_diagram.pdf}
    \caption{%
    The main components involved in our proposed disentangling method, $f_y$ (debiaser) and $h$ (discriminator). The debiaser is trained to produce encodings, $z_y$ of the data that are invariant to the source dataset and thereby the subgroups identifying it. In order to determine whether a bag of encodings originates from the training set or the deployment set, the discriminator performs an attention-weighted aggregation over the bag dimension to model interdependencies between the samples. In the case of Coloured MNIST where {\color{purple}purple} fours constitute the missing subgroup, the discriminator can identify an encoding of a bag from the training set by the absence of such samples so long as colour information is detectable in $z_y$, serving as an error signal for the debiaser.
    }%
    \label{fig:architecture}
\end{figure}%
% \end{figure}
%\end{wrapfigure}
% \hfill
% \begin{figure}
\begin{figure}[t]
    \centering
    % if you want to make changes to the diagram, it is located here:
    % https://drive.google.com/file/d/1M37m4rVj1EGrFJb9QzK32bs8swepiaZ1/view?usp=sharing
    \includegraphics[width=0.5\textwidth]{paper3/figures/fdm.pdf}
    \caption{%
    Illustration of distribution matching.
    In this example, the training set is lacking {\color{purple}purple} 4's.
    By enforcing the subspace $z_y$ to have the same distribution for both the training and deployment set, the model is encouraged to learn a representation that is invariant to colour (or $s$ in general.)
    For this to work, it is crucial that the bags be approximately balanced (perfect bags).
    }%
    \label{fig:matching-diagram}
% \end{subfigure}
% \caption{%
% Overview of the disentangling framework with a perfect (balanced) bag as an inductive bias.
% }%
% \label{fig:framework}
\end{figure}

@@

%Semi-supervised learning \cite{ChaSchZie06} can alleviate the issue of unfairness to the \emph{invisibles} by mixing labelled with unlabelled data, which is usually much cheaper to obtain. 
%
We propose to alleviate the issue of bias against missing subgroups by mixing labelled data with unlabelled data that is usually much cheaper to obtain \citep{ChaSchZie06}. 
%
In this paper, we refer to this set of \emph{unlabelled} data as the
deployment set\footnote{In our experiments, we report accuracy and bias metrics on another independent test set instead of on the unlabelled data that is available at training time.} $\gD_{dep} =\{(x_i)\}$.
%_{i=1}^{N^c}
%
This 
%context 
deployment set has overlap with all sources:
\begin{align}
  \gD_{dep} \cap \Omega_{y=y',s=s'} \neq \varnothing\quad\quad\forall y'\in\gY,\forall s'\in\gS~.
\end{align}
%The deployment set is much like the deployment set: 
Importantly, the deployment set has no information about class labels $y$ or the subgroup labels $s$.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=\columnwidth]{paper3/figures/celeba_dep_recon.png}
    \caption{%
    Visualization of our method's solutions for the CelebA dataset, with``smiling females'' as the missing subgroup.
    Column 1 shows the original images from $x$ from the deployment set of CelebA.
    Column 2 shows plain reconstructions generated from $x_\textit{recon}=g(f_y(x), f_s(x))$.
    Column 3 shows reconstruction with zeroed-out $z_s$: $g(f_y(x), 0)$, which effectively visualises $z_y$.
    Column 4 shows the result of an analogous process where $z_y$ was zeroed out instead.
    }%
    \label{fig:celeba-recons}
\end{figure}

@@

  \includegraphics[width=\columnwidth]{paper3/figures/cmnist_2v4_partial_acc.pdf}
  \includegraphics[width=\columnwidth]{paper3/figures/cmnist_2v4_partial_prr.pdf}
%   \includegraphics[width=\columnwidth]{paper3/figures/cmnist_2v4_partial_tpr.pdf}
%   \includegraphics[width=\columnwidth]{paper3/figures/cmnist_2v4_partial_tnr.pdf}

@@

  \includegraphics[width=\columnwidth]{paper3/figures/cmnist_2v4_miss_s_acc.pdf}
  \includegraphics[width=\columnwidth]{paper3/figures/cmnist_2v4_miss_s_prr.pdf}
%   \includegraphics[width=\columnwidth]{paper3/figures/cmnist_2v4_miss_s_alt_tpr.pdf}
%   \includegraphics[width=\columnwidth]{paper3/figures/cmnist_2v4_miss_s_alt_tnr.pdf}

@@

  \includegraphics[width=\columnwidth]{paper3/figures/cmnist_3dig_4miss_acc.pdf}
  \includegraphics[width=\columnwidth]{paper3/figures/cmnist_3dig_4miss_hgr.pdf}
%   \includegraphics[width=\columnwidth]{paper3/figures/cmnist_3dig_4miss_pr.pdf}
%   \includegraphics[width=\columnwidth]{paper3/figures/cmnist_3dig_4miss_tpr.pdf}
%   \includegraphics[width=\columnwidth]{paper3/figures/cmnist_3dig_4miss_tnr.pdf}

@@

  \includegraphics[width=\columnwidth]{paper3/figures/celeba_gender_smiling_acc.pdf}
  \includegraphics[width=\columnwidth]{paper3/figures/celeba_gender_smiling_prr.pdf}
%   \includegraphics[width=\columnwidth]{paper3/figures/celeba_gender_smiling_tpr.pdf}
%   \includegraphics[width=\columnwidth]{paper3/figures/celeba_gender_smiling_tnr.pdf}

@@

\paragraph{Constructing approximately perfect bags via clustering.}
We cluster the data points from the deployment set into $K=\textrm{dim}(\mathcal{Y})\cdot \textrm{dim}(\mathcal{S})$ number of clusters, i.e. the number of data sources $\Omega_{y,s}$. 
%, using a learned embedding $z$.
%
We use the k-means clustering algorithm, and a recently proposed method based on rank statistics \citep{HanRebEhrVedetal20}.
%
The cluster assignments can then be used to evenly stratify the deployment set into perfect batches, to be used by the subsequent disentangling phase.
%
% The objective for clustering can be written as a binary cross-entropy loss, induced by the following binary classification problem -- for any two data points in the deployment set to decide whether or not they belong to the same cluster:
% %
% \begin{align}
%  \mathcal{L}_{\textit{unsup}} &= -\frac{1}{N^2}\sum_{i,j} [s_{ij}\: \text{log} (C(z_i)^\top C(z_j)) \nonumber\\
%  &+ (1-s_{ij}) \: \text{log}(1-C(z_i)^\top C(z_j))].
%  \label{eq:clustering}
%  \end{align}
% %
% Here, $N^2$ is the amount of sample pairs,  $s_{ij}$ are binary labels capturing that similar data points belong to the same group, $s_{ij}=1$, while dissimilar ones do not, $s_{ij}=0$. 
% %
% Since the deployment set is unlabelled, $s_{ij}$ are unknown. 
% %
% We estimate $s_{ij}$ with the robust rank statistics approach (described below) and use the estimates as pseudo-labels in \eqref{eq:clustering}. 
% %
% The dot-product $C(z_i)^\top C(z_j)$ measures the similarity of the two data points $z_i$ and $z_j$ to score whether they share a cluster.
% %
% This is computed on the embedding of the softmax classifier $C: \mathcal{Z} \rightarrow \{1,\ldots, K\}$. 
% %
% In addition to this unsupervised training, the classifier $C$ is also trained on the labelled data using the standard cross-entropy loss.  
% %We train the classifier $C$ on the training data using a standard cross-entropy loss and the labelled sources $\Omega_{y,s}$. 
% %
% This way, the data points in the deployment set that are similar to the known demographics are assigned to the associated clusters, while the dissimilar data points form the new clusters corresponding to the invisible groups.

% The method of rank statistics \citep{YagStrRosLin11,DeaRuzSegShletal13,HanRebEhrVedetal20} avoids comparing the vectors $z_i$ and $z_j$ directly, and instead ranks the values in $z_i$ and in $z_j$ by their magnitudes, and compares the top-$U$ rankings (we found $U = 5$ to perform well in our experiments).
% %
% If the rankings obtained for two unlabelled data points $z_i$ and $z_j$ are the same, they are assigned to the same class, $s_{ij}=1$. 
% %
As a result of clustering, the data points in the deployment set $\gD_\mathit{dep}$ are labelled with cluster assignments 
$\gD_\mathit{dep}=\{(x_i, c_i)\}$, $c_i = C(z_i)$.
%
We balance $\gD_\mathit{dep}$ so that all clusters have equal size to form a perfect bag
% (see fig. \ref{fig:teaser})
, and use it as a supervision signal for the disentangling step.
%
%Note that our method of constructing the perfect dataset based on clustering does not require solving the linear assignment problem of cluster-source association. This is particularly suitable for exploring the setting where invisible demographics are formed of multiple sources.

@@

\input{paper3/appendix.tex}
